Choose a License

### 2024-09-18 choose_license.qmd
Create separate section on dual-licensing

"Choosing a License" involves:

1. Recording the copyright notices and licenses for files by others (for example, in a file called `LICENSE.txt`) and following all their additional requirements.
2. In the simplest case, choosing two licenses (one for code, one for everything else) which to apply  simultaneously to all own files (for example, again by recording the license in the file `LICENSE.txt`). Recipients are then free to choose under which of the two licenses to use the material.

@@

@@

If a Quarto document contains both R code and writing, one may wonder what type of license to apply. It is possible to make such a file simultaneously available under two licenses, at the choice of the recipient: Either under a specified software license, or under a Creative Commons license^[that is, a license for writing, image, audio, and video]. This is called dual-licensing and makes it easier to reuse both rendered PDF as well as the code. For example, one could write:

> The Quarto files in this project are made available under CC0\ 1.0 or (at your option) under the terms of the Apache Software License\ 2.0.

It is disputed whether software that uses an R package under AGPLv3 or GPLv3^[a variant of the AGPLv3 that does not cover software running as a service] can only be published under a GPL-compatible license -- or even has to be published under the same license. Posit, the company behind RStudio, does not believe that to be the case [see also @Wickham2023PackagesLicense]. You can learn which license an installed package uses via `packageDescription("<PACKAGE_NAME>", fields = "License")`. And to identify which licenses are being used by the R packages your project depends on, you can use the following code:

@@

Having selected the licenses of your choice -- again, you might need multiple ones depending on the types of works your project contains --, we encourage you to read through the full license text (or at least a legal summary) to understand their effect. Then, you can record the license of existing content and apply the licenses of your own contributions. Mostly, this just means indicating which license applies to which file or folder, usually in the project's README (among other places), whose creation will be discussed "[later](make_readme.qmd)".

@@

In the simplest case, one just creates a file called `LICENSE.txt` where the full text of the license is copied verbatim. This is a practice propagated by GitHub, which provides instructions for many licenses via [ChooseALicense.com](https://choosealicense.com/licenses/). However, if the project is not completely covered by one single license, this practice may become unwieldy. For example, if a project contains different types of works by different authors, the `LICENSE.txt` needs to detail which file is covered by which license(s), along with any copyright notices.

>>>>

_Choosing a license_ involves:

1. Recording the copyright notices and licenses for content by others (for example, in a file called `LICENSE.txt`) and following all their additional requirements.
2. In the simplest case, choosing two licenses (one suitable for code, one for everything else) which to apply  simultaneously to all own content (for example, again by recording the license in the file `LICENSE.txt`). Recipients are then free to choose under which of the two licenses to use the material.

@@

::: {#tip-dual-licensing .callout-tip}
#### Dual-licensing

Sometimes, the type of a work is not obvious. For example, a Quarto document...

- ...contains both R code and writing, and
- ...may be distributed in the source format or as rendered document, possibly including images.

One may wonder which license to apply in this case, because Creative Commons licenses are not recommended for source code^[because (among other reasons) they explicitly disclaim any conveyance of patent rights] and applying software licenses to PDFs or images can lead to confusion or nuisance.^[because they often require to keep the license notice intact]

One solution is to make such a work simultaneously available under two licenses, at the choice of the recipient: Either under a specified software license, or under a Creative Commons license^[that is, a license for writing, image, audio, and video]. This is called dual-licensing and makes it easier to reuse both the rendered document as well as the code. For example, one could write:

> The Quarto files in this project are made available under CC0\ 1.0 or (at your option) under the terms of the Apache Software License\ 2.0.
:::

@@

It is disputed whether software that uses an R package under AGPLv3 or GPLv3^[a variant of the AGPLv3 that does not cover software running as a service] can only be published under a GPL-compatible license -- or even has to be published under the same license. Posit, the company behind RStudio, does not believe that to be the case [see also @Wickham2023PackagesLicense].

You can learn which license an installed package uses via `packageDescription("<PACKAGE_NAME>", fields = "License")`. And to identify which licenses are being used by the R packages your project depends on, you can use the following code:

@@

Having selected the licenses of your choice -- again, you might need multiple ones depending on the types of works your project contains --, we encourage you to read through the full license text (or at least a legal summary) to understand their effect. Then, you can record the license of existing content and apply the licenses of your own contributions. Mostly, this just means indicating which license applies to which file or folder, usually in the project's README (among other places), whose creation will be discussed [later](make_readme.qmd).

@@

In the simplest case, one just creates a file called `LICENSE.txt` where the full text of the license is copied verbatim. This is a practice propagated by GitHub, which provides instructions for and comparisons of many licenses via [ChooseALicense.com](https://choosealicense.com/licenses/). However, if the project is not completely covered by one single license, this practice may become unwieldy. For example, if a project contains different types of works by different authors, the `LICENSE.txt` needs to detail which file is covered by which license(s), along with any copyright notices.
